Modelling the Retinotectal projection - can competition provide a stopping mechanism?

### 2021-07-16 paper/paper.tex
WIP, esp on interactions - added Ihist

Because there is convincing evidence that EphA and EphB receptors are
expressed in exponentially increasing
patterns \citep{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004},
we use an exponential form for retinal receptor expressions, in common with
other modelling
studies~\citep{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
%
The simulated tectum expresses the same types of receptors and ligands, also
in orthogonal pairs of gradients. In this work, we did not consider the effect
of reverse signalling from tectal receptors to RGC ligands so we only modelled
tectal ligand expression, $L$. Although several studies model tectal ligand
expression with exponential functions \citep{koulakov_stochastic_2004}, the
experimental evidence for ligand expression is more ambiguous (Supp. Fig
X). Correspondingly, we leave open the possibility that tectal ligand
expression may be modelled by exponential, linear or logarithmic functions.

@@

In this model, the signal transmitted when a ligand binds to a receptor on a
growth cone (forward signalling) can lead to one of two effects; the
signal may induce an attraction towards the ligand-expressing region or a
repulsion away from it.
%
For axon-tectum interactions, we implemented attraction as a climbing of the
tectal ligand expression gradient and repulsion as gradient descent.
%
To determine the strength of the effect we assumed a purely linear receptor
binding model, and set the chemotactic movement vector of the growth cone $b$ at
location $\mathbf{x}_b$ on the tectum to be
%
\begin{equation}
\mathbf{G}_b = \sum_i^N F_i\,r_{i,b} \nabla L_i(\mathbf{x}_b)
\end{equation}
%
where $r_{i,b}$ is the receptor expression on axon branch $b$ for
ligand-receptor pair $i$, $\nabla L_i$ is the gradient of expression of ligand
$i$ on the tectum and $F_i$ denotes the direction of the interaction induced
when a molecule of ligand $i$ binds to a receptor $i$ molecule. $F_i$ takes
the value -1 for a repulsive interaction or 1 for an attractive interaction.

@@

For axon-axon interactions, repulsion causes a movement of branch $b$ along a
unit vector, $\hat{\mathbf{x}}_{kb}$, from branch $k$ to branch $b$ causing
them to move further apart; attraction causes the opposite movement. The
interaction, $\mathbf{I}_b$, acting on axon branch $b$, is given by
%
\begin{equation}
\mathbf{I}_b = \frac{1}{|B_b|} \sum_k \hat{\mathbf{x}}_{kb}\,Q(d_{kb})
\end{equation}
%
where $Q(d_{kb})$ is the signalling strength between two growth cones of
radius $r$ a distance $d_{kb}$ from one another, given by
%
\begin{equation}
Q(d_{kb}) = \begin{cases}
     \sum_i^N F_i\,r_{i,b}\,l_{i,k}    & d_{kb} \leq 2r \\
     0 & d_{kb} > 2r
     \end{cases}
\end{equation}
%
The sign of $Q$ (which is dependent on the values of $F_i$) determines whether
the interaction, $\mathbf{I}_b$, is repulsive or attractive. $l_{i,k}$ is the
expression of ligand $i$ on branch $k$.

>>>>

Because there is convincing evidence that EphA and EphB receptors are
expressed in exponentially increasing
patterns \citep{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004},
we use an exponential form for retinal receptor expressions, in common with
other modelling
studies~\citep{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
%
The simulated tectum expresses the same types of receptors and ligands, also
in orthogonal pairs of gradients. In this work, we did not consider the effect
of reverse signalling from tectal receptors to RGC ligands so we only modelled
tectal ligand expression, $L$. Although several studies model tectal ligand
expression with exponential functions \citep{koulakov_stochastic_2004}, the
experimental evidence for ligand expression is more ambiguous (Supp. Fig
\textbf{X}). Correspondingly, we leave open the possibility that tectal ligand
expression may be modelled by exponential, linear or logarithmic functions.

@@

In this model, the signal transmitted when a ligand binds to a receptor on a
branch (forward signalling) can lead to one of two effects; the signal may
induce an attraction towards the ligand-expressing region or a repulsion away
from it.
%
For axon-tectum interactions, we implemented attraction as a climbing of the
tectal ligand expression gradient and repulsion as gradient descent.
%
To determine the strength of the effect we assumed a purely linear receptor
binding model, and set the chemotactic movement vector of the branch $b$ at
location $\mathbf{x}_b$ on the tectum to be
%
\begin{equation}
\mathbf{G}_b = \sum_i^N F_i\,r_{i,b} \nabla L_i(\mathbf{x}_b)
\end{equation}
%
where $r_{i,b}$ is the receptor expression on branch $b$ for ligand-receptor
pair $i$, $\nabla L_i$ is the gradient of expression of ligand $i$ on the
tectum and $F_i$ denotes the direction of the interaction induced when a
molecule of ligand $i$ binds to a receptor $i$ molecule. $F_i$ takes the value
-1 for a repulsive interaction or 1 for an attractive interaction.

@@

For axon-axon interactions, repulsion causes a movement of branch $b$ along a
unit vector, $\hat{\mathbf{x}}_{kb}$, from branch $k$ to branch $b$ causing
them to move further apart; attraction causes the opposite movement. The
interaction, $\mathbf{I}_b$, acting on branch $b$, is given by
%
\begin{equation}
\mathbf{I}_b = \frac{1}{|B_b|} \sum_k \hat{\mathbf{x}}_{kb}\,Q(d_{kb})
\end{equation}
%
where $Q(d_{kb})$ is the signalling strength between two growth cones of
radius $r$ a distance $d_{kb}$ from one another, given by
%
\begin{equation}
Q(d_{kb}) = \begin{cases}
     \sum_i^N F_i\,r_{i,b}\,l_{i,k}    & d_{kb} \leq 2r \\
     0 & d_{kb} > 2r
     \end{cases}
\end{equation}
%
The sign of $Q$ (which is dependent on the values of $F_i$) determines whether
the interaction, $\mathbf{I}_b$, is repulsive or attractive. $l_{i,k}$ is the
expression of ligand $i$ on branch $k$.
